---
title: "Porto Seguro's Safe Driver Prediction"
author: "kimnewzealand"
date: "11 November 2017"
output:
  html_notebook:
    fig_height: 4
    highlight: pygments
    theme: spacelab
  pdf_document: default
---

## Background

In this Kaggle competition, [https://www.kaggle.com/c/porto-seguro-safe-driver-prediction](Porto Seguro's Safe Driver Prediction), we are looking to predict the probability that an auto insurance policy holder files a claim. 

This notebook is an exploratory data analysis of the dataset.

* * *

### Setup

## Load packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, messages=FALSE}
library(data.table)
library(knitr)
library(tidyverse)
library(GGally)
library(stringr)
library(anchors)
library(corrplot)
```

```{r sessioninfo}
sessionInfo()
```


## Load Data

The first step will be to load the test data file and train data file. These have been downloaded to a local Kaggle folder offline and unzipped from the 7z file format using winzip.

We will use the data.table R package designed for large datasets.

```{r loaddata,include =FALSE}
## Using fread from data.table package to load work with large datasets
setwd("~/Kaggle/Porto")
train <- fread("./train.csv",colClasses = "numeric")
test <- fread("./test.csv",colClasses = "numeric")
```

1.2 Overview

In the dataset with `r dim(train)[1]` rows and `r dim(train)[2]` columns each row corresponds to a policy holder, and the _target_ column reflects that a claim was filed (1) or not (0).

_The following information is available from the Kaggle discussion forum: _ 

Features that belong to similar groupings are tagged as such in the feature names.

-"ind" is related to individual or driver,   
-"reg" is related to quality of life in a certain region,  
-"car" is related to car itself,  
-"calc" is an calculated feature.

In addition, feature names include the postfix _bin_ to indicate binary features and _cat_ to indicate categorical features. 

Features without these designations are either continuous or ordinal. 

Values of -1 indicate that the feature was missing from the observation. 

There are `r sum(is.na(train))` missing values or `r sum(is.na(train))/(nrow(train)*ncol(train))*100`  in the train dataset and `r sum(is.na(test))` missing values or `r sum(is.na(test))/(nrow(test)*ncol(test))*100` missing values in the test dataset.

* * *

### Part 1: Exploratory data analysis

1.1 Target Variable

Let's take a quick look at the target variable.

```{r summary}
head(train)
```

Take a look at the _target_ label variable as a histogram.

```{r targethistogram}
train %>% 
      ggplot(aes(as.factor(target),fill=as.factor(target)),guide=FALSE) + 
      geom_histogram(stat="count")+ 
      scale_y_continuous(name="Count",labels = scales::comma,limits = c(0,800000)) +
      xlab("Target") +
      ggtitle("Histogram of Target Variable") 

table(train$target)
```

We can see that a the number of claims as a proportion of total `r round(table(train$target)[2]/(table(train$target)[1]+table(train$target)[2]),2)` is very small. 

1.2 Pairs Plots of Feature Variables

We will be using the ggpairs function from the [GGally](https://cran.r-project.org/web/packages/GGally/index.html) package to pairs plot the categorical, binary and numerical variables. We will plot on a sample of the data as the train dataset is large and it will be time consuming to run the ggpairs on all data.

```{r trainsample}
# Random subset of train data
set.seed(123)
trainsample <- sample_n(train, 10000)
```

Plot the train data categorical variables.

```{r catggpairs}
# Plot train data categorical variables
ggpairs(trainsample[,c("target","ps_ind_02_cat","ps_ind_04_cat","ps_ind_05_cat")], upper = list(continuous = "points", combo = "box"), lower = list(continuous = "points", combo = "box"))

ggpairs(trainsample[,c("target","ps_car_01_cat","ps_car_02_cat","ps_car_03_cat","ps_car_04_cat","ps_car_05_cat","ps_car_06_cat")], upper = list(continuous = "points", combo = "box"), lower = list(continuous = "points", combo = "box"))

ggpairs(trainsample[,c("target","ps_car_07_cat","ps_car_08_cat","ps_car_09_cat","ps_car_10_cat","ps_car_11_cat")], upper = list(continuous = "points", combo = "box"), lower = list(continuous = "points", combo = "box"))
```


Plot the train data numerical variables. 

```{r numggpairs}
# Plot train data numerical variables
ggpairs(trainsample %>% 
              select(target,ps_ind_01,ps_ind_03,ps_ind_14,ps_ind_15))
ggpairs(trainsample %>% 
              select(target,ps_reg_01,ps_reg_02,ps_reg_03))
ggpairs(trainsample %>% 
              select(target,ps_car_11,ps_car_12,ps_car_13,ps_car_14,ps_car_15))
ggpairs(trainsample %>% 
              select(target,ps_calc_01, ps_calc_02, ps_calc_03, ps_calc_04, ps_calc_05, ps_calc_06, ps_calc_07, ps_calc_08, ps_calc_09))
```

Plot the train data binary variables.

```{r binggpairs}
# Plot train data binary variables
ggpairs(trainsample %>% 
              select(target,ps_ind_06_bin,ps_ind_07_bin,ps_ind_08_bin,ps_ind_09_bin,ps_ind_10_bin,ps_ind_11_bin,ps_ind_12_bin,ps_ind_13_bin,ps_ind_16_bin,ps_ind_17_bin,ps_ind_18_bin))
ggpairs(trainsample %>% 
              select(target,ps_calc_15_bin,ps_calc_16_bin,ps_calc_17_bin,ps_calc_18_bin,ps_calc_19_bin,ps_calc_20_bin))
```

1.3 Correlation Plots

We will use the [corrplot](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html) package to produce a correlation matrix, ignoring the calculated variables.

```{r}
train %>%
  dplyr::select(-starts_with("ps_calc")) %>%
  mutate_at(vars(ends_with("cat")), funs(as.integer)) %>%
  mutate_at(vars(ends_with("bin")), funs(as.integer)) %>%
  mutate(target = as.integer(target)) %>%
  cor( method = "spearman") %>%
      corrplot(type="lower",method="square", tl.col = "black",  diag=FALSE)
```

### Part 2: Data Cleaning

Let's clean the train and test datasets variable by variable based on the exploratory data analysis.

2.1 Remove ID

```{r removeid}
# Remove ID feature from train as this identifier is not needed in the training. The id will be left in the test set for the predictions.
train <- train %>% dplyr::select(-id)
```

2.2 Missing values

Next we will extract and view the variables in more detail that have value -1, which represent the missing values in the dataset. With dplyr we can call functions from different R packages directly inside the dplyr functions. We will use the [stringr](https://cran.r-project.org/web/packages/stringr/vignettes/stringr.html) R package with dplyr to view a summary of the -1's. Then we will then use the [anchors](https://cran.r-project.org/web/packages/anchors/index.html) R package with the replace.value function to replace the -1's with the column means.


```{r recode}
# Use the base summary function for result summaries not dplyr. This will provide us with the ranges of the variables including the minimums
s <- summary(train)  
#  Extract and view the min values that have -1 from the summary we just created use str_detect from stringr package. 
s %>% 
      data.frame() %>% 
      filter(str_detect(Freq,"-1")) %>% 
      filter(str_detect(Freq,"Min")) %>%
      dplyr::select(-1)

# Replace the -1 ie missing values in the numeric columns with the mean of the respective column using recode, first replacing the -1 with NAs otherwise the -1 will distort the calculated mean

# Find the index of the columns names that contain cat
indx <- grepl('cat', colnames(train))
# First convert the categorical columnds to NA
train <- replace.value( train, colnames(train)[indx], from=-1, to=NA, verbose = FALSE)
# Calculate the rounded mean for a sanity check
round(mean(train$ps_car_03_cat,na.rm = TRUE)) # starts with -1,-1,-1,0,-1
round(mean(train$ps_car_05_cat,na.rm = TRUE)) # starts with 1,-1,-1,1
# Create the columns mean functions for catageorical means
roundmean <- function(x) {replace(x, is.na(x), round(mean(x, na.rm = TRUE))) } 
# Replace the NAs with the roundmean
train <- as.data.frame(apply(train, 2, roundmean))
# Sanity check against the rounded means calculated above
train$ps_car_03_cat[2]
train$ps_car_05_cat[2]
# Next convert the numericall columns to NA
train <- replace.value( train, colnames(train)[-indx], from=-1, to=NA, verbose = FALSE)
# Calculate mean before recoding
mean(train$ps_reg_03,na.rm = TRUE) # row 3
# Create the columns mean functions , one for mean of the continuous numerical columns
justmean <- function(x) {replace(x, is.na(x), mean(x, na.rm = TRUE)) }
# Replace the NAs with the justmean
train <- as.data.frame(apply(train, 2, justmean))
#Sanity check this is same mean as calculated above 
train$ps_reg_03[3]
# Sanity check that we have cleaned up all -1's, the result should be empty
colsum <- colSums(train=="-1") 
colsum[colsum>0]
```

